---
title: "scaled_density"
date: "2023-02-17"
output: html_document
---
calculate wood specific gravity (using ONLY undamaged cookies)

  1. species-level wood specific gravity: take all undamaged cookies for a given species, find the average wood specific gravity. 
  2. tree-level wood specific gravity: take all undamaged cookies for a given species, find the relationship between size and wood specific gravity. then apply the relationship (by species) to each tree cylinder model (using size to predict specific gravity).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# SET WORKING DIRECTORY HERE
setwd("path/to/your/desired/working_directory")
```

libraries 
```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
```

data load and format
```{r}
# read in data - cookie
load("MEE_repo/R/data_out/cookie.RData")

# load cookie main data 
load("MEE_repo/R/data_out/main.RData")
main <- main[c("field_ID", "tree", "firebreak", "cutpoint", "species", "perc_water_content", "density_fresh_disp_gcm3", "density_dry_disp_gcm3", "diam_1", "diam_2")]

# create new ID column to merge tree ID and cutpoint 
cookie$drill_id <- paste(cookie$tree_ID,cookie$position, sep = "_")
cookie <- cookie %>% relocate(drill_id, .before = tree_ID)

# merge cookie damage data with master (contains wood density)
by <- c("field_ID", "species")
df <- full_join(cookie, main, by = by)

# remove C cookies (no information on density or WC, so use as a conditional)
df %>% drop_na(perc_water_content) -> df

# pull out and process terminal branches seperately
term <- df[is.na(df$tree_ID),]
all <- df[!is.na(df$tree_ID),]

# fix terminal branch information (fill in missing pieces)
term$position <- term$cutpoint
term$tree.x <- term$tree.y
term$tree_ID <- paste(term$tree.x, term$firebreak, sep = "_") # paste tree.x with firebreak 
term$drill_id <- paste(term$tree_ID, term$position, sep = "_") # paste tree_ID with position 
term$diameter_cm <- (term$diam_1 + term$diam_2) / 2 # find average diameter of terminal branches 

# enter no damage for terminal branches
term %>% 
  mutate(damaged = ifelse(is.na(damaged), "n", damaged)) -> term

# clean up NAs (EXTRA3-BR2 does not have a cookie photo, therefore no diameter)
term %>% drop_na(diameter_cm) -> term

# merge terminal branches back to all other cookies - working df for wood density analysis
wd <- rbind(all, term)

# change df structure to create binary categorical damage variable 
wd$damaged <-  as.factor(wd$damaged)

# subset working df that only includes undamaged cookies (to avoid calculating biased wood density from damaaged cookies)
wd_ND <- wd[!(wd$damaged == "y"),]
```

calculate species level wood density 
```{r}
# without terminal branches (these are over-represented because they all have no damage)
wd_ND_NT <- wd_ND[!(wd_ND$position == "TERM" | wd_ND$cutpoint == "TERM-A" | wd_ND$cutpoint =="TERM-B"),]
wd_ND_NT %>%                               
  group_by(species) %>% 
  summarize(min = min(density_dry_disp_gcm3),
            q1 = quantile(density_dry_disp_gcm3, 0.25),
            median = median(density_dry_disp_gcm3),
            mean = mean(density_dry_disp_gcm3),
            q3 = quantile(density_dry_disp_gcm3, 0.75),
            max = max(density_dry_disp_gcm3), 
            n = n(), 
            sd = sd(density_dry_disp_gcm3)) -> density_summary_sp_noterm

# match to ply model by species 
trees <- read.csv("MEE_repo/data/Iron_Range_Stem_Damage_master_Mar_02_23.xlsx - all (1).csv")

trees_wd <- trees[c("tree_id", "species")]
trees_species_wd <- left_join(trees_wd, density_summary_sp_noterm, by = "species")

field_wd_species <- trees_species_wd[c("tree_id", "species", "mean", "sd")]
names(field_wd_species)[names(field_wd_species) == 'mean'] <- 'wd_species_level_field'
```
