---
title: "Untitled"
date: "2023-05-31"
output: html_document
---

library
```{r}
library(dplyr)
library(viridis)
library(rcompanion)
library(glmmTMB)
library(DescTools)
library(lmtest)
library(dplyr)
library(ggeffects)
library(forcats)
library(tidyverse)
library(lme4)
library(ggpubr)
library(ggpmisc)
library(easystats)
library(car)
library(glmmTMB)
library(report) 
library(ggplot2)
```

load data
```{r}
# SET WORKING DIRECTORY HERE
setwd("path/to/your/desired/working_directory")
# all cookies
load("MEE_repo/R/data_out/cookie.RData")
# damaged cookies
load("MEE_repo/R/data_out/cookie_d.RData")
# cookie main sheet
load("MEE_repo/R/data_out/main.RData")
# tree damage
load("MEE_repo/R/data_out/tree_level_damage.RData")
# tree biomass main sheet (trees)
trees <- read.csv("MEE_repo/R/data_out/all_trees_ex2fixed.csv")
# damaged cookies with height in tree
load("MEE_repo/R/data_out/cookie_heights_added_out.RData")
# tree biomass
data <- as.data.frame(read.csv("all_trees_Abbey_June.csv"))
```

Figure 4. total AGB plot
```{r}
# Paul2013 scatterplot
(paul2013 <- ggplot(data, aes(x = obs_ABG_dry_kg, y = paul2013)) + 
   geom_abline(alpha = 0.5, linetype = 4) + 
   geom_point(alpha = 0.3) +
   theme_classic() +
   xlim(0,3100) +
   ylim(0,3100) +
   xlab("AGB (kg)") +
   ylab("Paul 2013 ASM (kg)")) +
   stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth",
              colour="black")

(paul2013_inset <- ggplot(data, aes(x = obs_ABG_dry_kg, y = paul2013)) + 
   geom_abline(alpha = 0.5, linetype = 4) + 
   geom_point(alpha = 0.3) +
   theme_classic() +
   xlim(0,300) +
   ylim(0,300) +
   xlab("AGB (kg)") +
   ylab("Paul 2013 ASM (kg)")) +
   stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth",
              colour="black")

#Chave2014 scatterplot field densities
(chave2014 <- ggplot(data, aes(x = obs_ABG_dry_kg, y = chave_F)) + 
   geom_abline(alpha = 0.5, linetype = 4) + 
   geom_point(alpha = 0.3) +
   theme_classic() +
   xlim(0,3100) +
   ylim(0,3100) +
   xlab("AGB (kg)") +
   ylab("Chave 2014 ASM (kg)")) +
   stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth",
              colour="black")

(chave2014_inset <- ggplot(data, aes(x = obs_ABG_dry_kg, y = chave_F)) + 
   geom_abline(alpha = 0.5, linetype = 4) + 
   geom_point(alpha = 0.3) +
   theme_classic() +
   xlim(0,300) +
   ylim(0,300) +
   xlab("AGB (kg)") +
   ylab("Chave 2014 ASM (kg)")) +
   stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth",
              colour="black")

# TLS scatterplot field densities 
(TLS <- ggplot(data, aes(x = obs_ABG_dry_kg, y = TLS_F)) + 
   geom_abline(alpha = 0.5, linetype = 4) + 
   geom_point(alpha = 0.3) +
   theme_classic() +
   xlim(0,3100) +
   ylim(0,3100) +
   xlab("AGB (kg)") +
   ylab("TLS modelled weights (kg)")) +
   stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth",
              colour="black")

(TLS_inset <- ggplot(data, aes(x = obs_ABG_dry_kg, y = TLS_F)) + 
   geom_abline(alpha = 0.5, linetype = 4) + 
   geom_point(alpha = 0.3) +
   theme_classic() +
   xlim(0,300) +
   ylim(0,300) +
   xlab("AGB (kg)") +
   ylab("TLS modelled weights (kg)")) +
   stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth",
              colour="black")

# comparison of all ASM/TLS to baseline value of observed biomass 
trees_total <- trees[c("paul2016", "paul2013", "brown", "williams", "chave_F", "TLS_F","chave_L", "TLS_L","obs_ABG_dry_kg")]

sums <- data.frame(colSums(trees_total, na.rm = TRUE))
sums$mod <- rownames(sums)
colnames(sums)[1] = "estimate"
sums$baseline_difference <- sums$estimate - sums[9,1]
sums$percent_difference <- (sums$baseline_difference/(sums[9,1]))*100

# plot of ASM/TLS difference from baseline (field-measured) as %
AGBb <- ggplot(sums, aes(x = mod, y = percent_difference,  fill = mod)) + 
  geom_bar(stat = "identity")+
  ylab("% difference from field-measured AGB")+
  xlab("")+
  scale_x_discrete(limits=c("chave_F", "TLS_F", "paul2016", "paul2013", "brown", "williams", "chave_L", "TLS_L"),
                   labels=c("Chave (field p)", "TLS (field p)","Paul 2016", "Paul 2013", "Brown 1997",  "Williams 2005", "Chave (DB p)", "TLS (DB p)"))+
  theme_classic()+
  theme(legend.position = "Spectral")+ 
  geom_hline(yintercept=0, linetype="dashed", color = "red")+
  ggtitle("b)")+
  scale_fill_viridis(discrete = TRUE)

```

Table 1. model regressions (ASM and TLS)
```{r}
#Williams
reg <- lm(formula = williams ~ obs_ABG_dry_kg,
   data=data)
summary(reg)

#Brown
reg <- lm(formula = brown ~ obs_ABG_dry_kg,
   data=data)
summary(reg)

#Paul2013
reg <- lm(formula = paul2013 ~ obs_ABG_dry_kg,
   data=data)
summary(reg)

# Paul2016
reg <- lm(formula = paul2016 ~ obs_ABG_dry_kg,
   data=data)
summary(reg)

#Chave2014
reg <- lm(formula = chave_F ~ obs_ABG_dry_kg,
   data=data)
summary(reg)

#TLS
reg <- lm(formula = TLS_F ~ obs_ABG_dry_kg,
   data=data)
summary(reg)
```

Figure 5. damage throughout tree plot
```{r}
# convert to proportion
cookie_d$all_dam_prop <- cookie_d$all_dam/(100) 

# reorder positions to correct levels
cookie_d <- cookie_d %>% mutate(position = fct_relevel(position, c("S", "C", "A", "BR2", "BR3", "BR4", "BR5")))%>%
  mutate(nondmg = 1 - all_dam_prop)

cookie_heights_added_out <- cookie_heights_added_out %>%
  mutate(nondmg_area = cross_section_area - area_dam_cm2)

# fix one negative height 
cookie_heights_added_out$height_m <- ifelse(cookie_heights_added_out$height_m < 0, 0, cookie_heights_added_out$height_m)
cookie_heights_added_out$relative_height_prop <- cookie_heights_added_out$relative_height/100
cookie_heights_added_out$relative_height_prop <- ifelse(cookie_heights_added_out$relative_height_prop < 0, 0, cookie_heights_added_out$relative_height_prop)

# model
log.model <-lmer(log(all_dam+1) ~ relative_height + (1|tree_ID), cookie_heights_added_out) # best 
summary(log.model)
Anova(log.model)

# visualization 
height_fig <- ggpredict(log.model, terms = c("relative_height")) %>% plot(add.data = TRUE)
height_fig <- height_fig+
  coord_flip()

predicted_height <- ggpredict(log.model,terms = c("relative_height"),)
ggplot(cookie_heights_added_out, aes(x = relative_height, y = all_dam, color=position)) + 
  geom_jitter() +
  geom_line(aes(x=x, y=predicted), data=predicted_height, inherit.aes=FALSE)+
  theme_classic()+
  scale_color_viridis(limits=c("S", "C", "A", "BR2", "BR3", "BR4", "BR5"), labels = c("C1", "C2", "C3", "C4", "C5", "C6", "C7"), discrete = TRUE, direction = 1)+
  xlim(0,100)+
  coord_flip()
```

Figure 6. damage residual plot 
```{r}
# merge damage and trees 
dam_scaled_sub <- dam_scaled[c("ply_model", "mu_hol_scaled", "mu_rot_scaled", "mu_tot_scaled")]
trees_damage <- left_join(trees, dam_scaled_sub, by = "ply_model")
trees_damage$mu_hol_scaled[is.na(trees_damage$mu_hol_scaled)] <- 0
trees_damage$mu_rot_scaled[is.na(trees_damage$mu_rot_scaled)] <- 0
trees_damage$mu_tot_scaled[is.na(trees_damage$mu_tot_scaled)] <- 0
trees_damage$damaged <- ifelse(trees_damage$mu_tot_scaled>0, "Y", "N")
trees_damage$residuals <- trees_damage$TLS_F-trees_damage$obs_ABG_dry_kg

# TLS - plot relationship between damage % and residuals
dam_TLS <- ggplot(trees_damage[trees_damage$damaged=="Y",], aes(x = mu_tot_scaled, y = residuals/obs_ABG_dry_kg, size = dbh_observed_cm_tidy))+
  geom_point(alpha = 0.25)+
 # geom_smooth(method = "lm" )+
  geom_hline(yintercept=0, linetype="dashed", color = "red")+
  theme_classic()+
  xlab("Tree-level damage (%)")+
  ylab("Difference between TLS and field measured AGB (kg), normalized for size")+
  ggtitle("b)")+
  ylim(-1.5, 3)

trees_damage$residuals_scaled <- trees_damage$residuals/trees_damage$obs_ABG_dry_kg
mod.tls <- lm((residuals_scaled)~mu_tot_scaled*dbh_observed_cm_tidy, data = trees_damage[trees_damage$damaged=="Y",])
summary(mod.tls)

dam_tls <- ggpredict(mod.tls, terms = c("mu_tot_scaled", "dbh_observed_cm_tidy")) %>% plot(add.data = TRUE) # with raw data color by size
dam_tls <- dam_tls + 
  ylim(-1.5, 3)+
  ggtitle("a)")

sm <- "#fce725"
med <- '#20908c'
lg <- '#440153'
v <- c(13, 27, 40)
predicted_residuals_tls <- ggpredict(mod.tls,terms = c("mu_tot_scaled", "dbh_observed_cm_tidy [v]"),)

fig6a <-ggplot(trees_damage[trees_damage$damaged=="Y",], aes(x = mu_tot_scaled, y = residuals/obs_ABG_dry_kg, size = dbh_observed_cm_tidy)) + 
  geom_jitter(alpha = 0.25) +
  geom_line(aes(x=x, y=predicted), data=filter(predicted_residuals_tls, group=='13'), 
            inherit.aes=FALSE, colour=sm) + 
  geom_ribbon(aes(x=x, ymin=conf.low, ymax=conf.high), data=filter(predicted_residuals_tls, group=='13'), 
              inherit.aes=FALSE, alpha=0.2, fill = sm)+
  geom_line(aes(x=x, y=predicted), data=filter(predicted_residuals_tls, group=='27'), 
            inherit.aes=FALSE, colour=med) + 
  geom_ribbon(aes(x=x, ymin=conf.low, ymax=conf.high), data=filter(predicted_residuals_tls, group=='27'), 
              inherit.aes=FALSE, alpha=0.2, fill = med)+
  geom_line(aes(x=x, y=predicted), data=filter(predicted_residuals_tls, group=='40'), 
            inherit.aes=FALSE, colour=lg) + 
  geom_ribbon(aes(x=x, ymin=conf.low, ymax=conf.high), data=filter(predicted_residuals_tls, group=='40'), 
              inherit.aes=FALSE, alpha=0.2, fill = lg)+
  theme_classic()+
  ylim(-1.5, 3)+
  geom_hline(yintercept=0, linetype="dashed", color = "red")+
  ggtitle("a)")

# ASM - plot relationship between damage % and residuals
# calculate ASM residuals
trees_damage$residuals_ASM <- trees_damage$chave_F-trees_damage$obs_ABG_dry_kg

dam_ASM <- ggplot(trees_damage[trees_damage$damaged=="Y",], aes(x = mu_tot_scaled, y = residuals_ASM/obs_ABG_dry_kg, size = dbh_observed_cm_tidy ))+
  geom_point(alpha = 0.25)+
  geom_hline(yintercept=0, linetype="dashed", color = "red")+
  theme_classic()+
  xlab("Tree-level damage (%)")+
  ylab("Difference between ASM and field measured AGB (kg), normalized for size")+
  ggtitle("d")+
  ylim(-1.5, 3)

mod.asm <- lm((residuals_ASM/obs_ABG_dry_kg)~mu_tot_scaled*dbh_observed_cm_tidy, data = trees_damage[trees_damage$damaged=="Y",])
summary(mod.asm)

# plot(mod.asm)
dam_asm <- ggpredict(mod.asm, terms = c("mu_tot_scaled", "dbh_observed_cm_tidy")) %>% plot(add.data = TRUE)

predicted_residuals_asm <- ggpredict(mod.asm,terms = c("mu_tot_scaled", "dbh_observed_cm_tidy [v]"),)

fig6b <- ggplot(trees_damage[trees_damage$damaged=="Y",], aes(x = mu_tot_scaled, y = residuals_ASM/obs_ABG_dry_kg, size = dbh_observed_cm_tidy)) + 
  geom_jitter(alpha = 0.25) +
  geom_line(aes(x=x, y=predicted), data=filter(predicted_residuals_asm, group=='13'), 
            inherit.aes=FALSE, colour=sm) + 
  geom_ribbon(aes(x=x, ymin=conf.low, ymax=conf.high), data=filter(predicted_residuals_asm, group=='13'), 
              inherit.aes=FALSE, alpha=0.2, fill = sm)+
  geom_line(aes(x=x, y=predicted), data=filter(predicted_residuals_asm, group=='27'), 
            inherit.aes=FALSE, colour=med) + 
  geom_ribbon(aes(x=x, ymin=conf.low, ymax=conf.high), data=filter(predicted_residuals_asm, group=='27'), 
              inherit.aes=FALSE, alpha=0.2, fill = med)+
  geom_line(aes(x=x, y=predicted), data=filter(predicted_residuals_asm, group=='40'), 
            inherit.aes=FALSE, colour=lg) + 
  geom_ribbon(aes(x=x, ymin=conf.low, ymax=conf.high), data=filter(predicted_residuals_asm, group=='40'), 
              inherit.aes=FALSE, alpha=0.2, fill = lg)+
  theme_classic()+
  ylim(-1.5, 3)+
  geom_hline(yintercept=0, linetype="dashed", color = "red")+
  ggtitle("b)")
  
ggarrange(fig6a,fig6b, ncol = 2, legend = "right", common.legend = TRUE)
```

Supplementary Figure 1. all model biomass figures (excluding Paul2013, Chave and TLS using field wd from fig1)
```{r}
#Williams2005 scatterplot
(williams <- ggplot(data, aes(x = obs_ABG_dry_kg, y = williams)) + 
   geom_abline(alpha = 0.5, linetype = 4) + 
   geom_point(alpha = 0.3) +
   theme_classic() +
   xlim(0,3100) +
   ylim(0,3100) +
   xlab("AGB (kg)") +
   ylab("Williams 2005 ASM (kg)")) +
   stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth",
              colour="black")

(williams_inset <- ggplot(data, aes(x = obs_ABG_dry_kg, y = williams)) + 
   geom_abline(alpha = 0.5, linetype = 4) + 
   geom_point(alpha = 0.3) +
   theme_classic() +
   xlim(0,300) +
   ylim(0,300) +
   xlab("AGB (kg)") +
   ylab("Williams 2005 ASM (kg)")) +
   stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth",
              colour="black")

#Brown scatterplot
(brown <- ggplot(data, aes(x = obs_ABG_dry_kg, y = brown)) + 
   geom_abline(alpha = 0.5, linetype = 4) + 
   geom_point(alpha = 0.3) +
   theme_classic() +
   xlim(0,3100) +
   ylim(0,3100) +
   xlab("AGB (kg)") +
   ylab("Brown 1997 ASM (kg)")) +
   stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth",
              colour="black")

(Brown_inset <- ggplot(data, aes(x = obs_ABG_dry_kg, y = brown)) + 
   geom_abline(alpha = 0.5, linetype = 4) + 
   geom_point(alpha = 0.3) +
   theme_classic() +
   xlim(0,300) +
   ylim(0,300) +
   xlab("AGB (kg)") +
   ylab("Brown 1997 ASM (kg)")) +
   stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth",
              colour="black")

#Paul2016 scatterplot
(paul16 <- ggplot(data, aes(x = obs_ABG_dry_kg, y = paul2016)) + 
   geom_abline(alpha = 0.5, linetype = 4) + 
   geom_point(alpha = 0.3) +
   theme_classic() +
   xlim(0,3100) +
   ylim(0,3100) +
   xlab("AGB (kg)") +
   ylab("Paul 2016 ASM (kg)")) +
   stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth",
              colour="black")

(paul16_inset <- ggplot(data, aes(x = obs_ABG_dry_kg, y = paul2016)) + 
   geom_abline(alpha = 0.5, linetype = 4) + 
   geom_point(alpha = 0.3) +
   theme_classic() +
   xlim(0,300) +
   ylim(0,300) +
   xlab("AGB (kg)") +
   ylab("Paul 2016 ASM (kg)")) +
   stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth",
              colour="black")

#Chave2014 scatterplot w/ published densities
(chave2014_L <- ggplot(data, aes(x = obs_ABG_dry_kg, y = chave_L)) + 
   geom_abline(alpha = 0.5, linetype = 4) + 
   geom_point(alpha = 0.3) +
   theme_classic() +
   xlim(0,3500) +
   ylim(0,3500) +
   xlab("AGB (kg)") +
   ylab("Chave 2014 ASM reference pwood (kg)")) +
   stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth",
              colour="black")

(chave2014_L_inset <- ggplot(data, aes(x = obs_ABG_dry_kg, y = chave_L)) + 
   geom_abline(alpha = 0.5, linetype = 4) + 
   geom_point(alpha = 0.3) +
   theme_classic() +
   xlim(0,300) +
   ylim(0,300) +
   xlab("AGB (kg)") +
   ylab("Chave 2014 ASM ref pwood (kg)")) +
   stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth",
              colour="black")

#TLS scatterplot w published densities
(TLS_L <- ggplot(data, aes(x = obs_ABG_dry_kg, y = TLS_L)) + 
   geom_abline(alpha = 0.5, linetype = 4) + 
   geom_point(alpha = 0.3) +
   theme_classic() +
   xlim(0,3500) +
   ylim(0,3500) +
   xlab("AGB (kg)") +
   ylab("TLS modelled weights ref pwood (kg)")) +
   stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth",
              colour="black")

(TLS_inset <- ggplot(data, aes(x = obs_ABG_dry_kg, y = TLS_L)) + 
   geom_abline(alpha = 0.5, linetype = 4) + 
   geom_point(alpha = 0.3) +
   theme_classic() +
   xlim(0,300) +
   ylim(0,300) +
   xlab("AGB (kg)") +
   ylab("TLS modelled weights ref pwood (kg)")) +
   stat_smooth(method = "lm",
              formula = y ~ x,
              geom = "smooth",
              colour="black")
```

Supplementary Figure 2. density scatterplot
```{r}
(densities <- ggplot(data, aes(x = dry_density_published, y = wd_species_level_field, color = species)) + 
   geom_abline(alpha = 0.5, linetype = 2) + 
   geom_point(size = 2.5, stroke = 0.3) +
   theme_classic() +
   xlim(0.4,1.05) +
   ylim(0.4,1.05) +
   xlab("Global Wood Density Database (Zanne et al. 2009)") +
   ylab("Species mean field specific gravity"))
```

Supplementary Figure 4. tree size distribution 
```{r}
# all trees 
alltree <- ggplot(trees, aes(dbh_observed_cm_tidy)) +
  geom_bar() +
  scale_x_binned()+ 
  theme_classic()+
  ylim(0,22)+
  xlab("DBH (cm)")+
  ylab("Number of Individuals")
  
# just damamged trees 
damtree <- ggplot(trees_damage[trees_damage$damaged=="Y",], aes(dbh_observed_cm_tidy)) +
  geom_bar() +
  scale_x_binned()+ 
  theme_classic()+
  ylim(0,22)+
  xlab("DBH (cm)")+
  ylab("Number of Individuals")
ggarrange(alltree, damtree)
```

Supplementary Figure 5. undamaged trees - size relationship with residuals
```{r}
aa <- ggplot(trees_damage[trees_damage$damaged=="N",], aes(x = dbh_observed_cm_tidy, y = residuals/obs_ABG_dry_kg)) +
  geom_point()+
  theme_classic()+
  xlab("DBH (cm)")+
  ylab("Difference between TLS and field measured AGB (kg), normalized for size")+
  ggtitle("a)")+
  ylim(0, 2)

ab <- ggplot(trees_damage[trees_damage$damaged=="N",], aes(x = dbh_observed_cm_tidy, y = residuals_ASM/obs_ABG_dry_kg)) +
  geom_point()+
  theme_classic()+
  xlab("DBH (cm)")+
  ylab("Difference between ASM and field measured AGB (kg), normalized for size")+
  ggtitle("b)")+
  ylim(0, 2)

ggarrange(aa, ab, nrow = 1)
```
